Comprehensive Algorithmic Resilience for \\Numeric Applications

### 2013-10-14 paper.tex
Updating abstract and intro for current paper angle.

\begin{abstract}
As High-Performance Computing (HPC) systems become larger and more complex and the feature sizes of their electronic components grow smaller, the systems become increasingly vulnerable to soft faults.
Since such faults may cause application crashes or may silently corrupt application results it is necessary to develop techniques to protect applications from such errors at a low cost in performance and power.
Despite extensive work on general-purpose resilience techniques, they are cost effective for protecting only certain system components, such as memories.
To provide comprehensive protection for the computations of HPC applications it is necessary to develop algorithmic resilience mechanisms that use the properties of algorithms to provide cost-effective protection.

\sui{
Algorithmic resilience exploiting the invariants of certain algorithms is an old technique that has been around for many years. The invariants can serve as
correctness unit tests during development and turned off during run-time in favor of speed, but under the context of the upcoming exascale reliability challenges,
they can usually serve as useful fault checkers. However, unless other issues including input protection, control information
and critical data structure protection are resolved will algorithmic invariants help detecting and correcting errors.
% can be very complex to deploy in applications because a comprehensive resilience strategy must address all the types of errors that an application may experience.
%This includes corruptions to numerical data as well as control information and data structures.
%Full protection requires the combination of multiple techniques including algorithmic error detection, crash detection, replication of key control information as well as support for restart when errors are detected.
%Further, it is necessary to configure the properties of the resilience mechanisms to make sure that they are cost-effectively meeting the user's accuracy and confidence targets.
Further, certain design choices in exploiting algorithmic resilience imply trade-offs for cost-effectiveness of the final resilience strategy.
This paper presents our case study of exploiting algorithmic invariants for constructing comprehensive fault tolerance mechanisms and balancing performance and result accuracy in the context of large-scale computing.
}
%This paper shows how all these issues can be resolved comprehensively in the context of three numeric applications.

@@

The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
While this goal is important, it ignores the full complexity of the problem since it focuses on only a subset of the errors \sui{(algorithmic correctness)} and 
does not consider combinations of different resilience techniques to efficiently and effectively address all the different types of faults.
In reality a comprehensive solution must incorporate a range of additional techniques such the replication of key variables, detection of memory violations and localized rollback on detected errors, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
%, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
%Fortunately, because real-world applications spend most of their execution is spent in a few application modules, most soft faults will occur in those modules rather than uniformly throughout the application.
%In scientific applications these modules are usually various types of libraries, including numeric solvers and domain-specific scientific packages.
%This indicates that if by focusing algorithmic resilience efforts on the key library routines, while leaving other modules unprotected or protected via more expensive mechanisms, it is possible to fully protect an application at a low cost.

This paper presents a comprehensive algorithmic resilience approach \sui{that combines existing checkpointing and algorithmic checker techniques that considers all of the above issues and demonstrate {\bf with fault injection experiments and statistical analyses} that such approach is able to significantly reduce the side effects of soft errors on parallel programs at a  slight performance cost.}
We demonstrate it on different types of numerical applications: the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to the routines where they spend the most time: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
We demonstrate via fault injection experiments that although none of these techniques can individually protect applications from soft faults, their combination is highly effective.
Further, we demonstrate how these techniques can be configured to offer the right level of protection at any given fault rate or input type.
More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
%\greg{Summarize real results}
%\sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09)}
%For the three considered scientific codes, we correct more than 50\% of segmentation faults and we reduce the error in the outputs by one order of magnitude. The overhead is always lower than 20\%.

>>>>

\begin{abstract}
As High-Performance Computing (HPC) systems become larger and more complex and the feature sizes of their electronic components grow smaller, the systems become increasingly vulnerable to soft faults.
Since such faults may cause application crashes or may silently corrupt application results it is necessary to develop techniques to protect applications from such errors at a low cost in performance and power.
\greg{Since general-purpose resilience mechanisms, such as replication, can be very expensive for general applications it is imperative to develop ways to tailor resilience techniques to the needs and properties of each application.}

\greg{Extensive ongoing work on algorithmic techniques to detect and tolerate errors in common scientific sub-routines and kernels has shown that such techniques can be very effective at protecting the numeric properties of these code regions.
However, the practical utility of these techniques has been limited by the fact that real scientific applications (i) are significantly more complex than these simple kernels and (i) are vulnerable to a wider range of errors than most algorithmic checkers focus on.
To make scientific applications resilient to soft errors it is necessary to comprehensively combine resilience techniques that detect all the major types of errors that can affect an application and provide techniques that can tolerate them.
Further, the techniques applied to a given application must be tailored to an application's needs and configured to provide the performance for each input, hardware error rate and application user's accuracy tolerance.
This paper presents a case study of the design of comprehensive resilience mechanisms for four(\textbf{???}) different applications.
In addition to exploring the space of mechanism design and configuration we project our results to massively parallel applications and show how application performance and resilience configuration choices vary with error frequency, application scale and accuracy goals.
}

@@

\greg{The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
While this work is crucial, it addresses only a portion of the full challenge of making scientific applications resilient to soft faults.
A comprehensive resilience strategy must address three areas.
First, it must deal all the possible application-level manifestations of soft faults including numeric errors (the focus of traditional algorithmic checkers), as well as corruptions in data structures and other application state and computations.
Second, it must address tolerance to soft faults in addition to their detection.
Finally, given the many ways in which resilience techniques can be configured (e.g. different detection tolerances for an algorithmic checker) it must provide ways to choose the best configuration for each system and for each user's performance and accuracy needs.}

\greg{
This paper presents a case study of how existing resilience techniques can be combined to make a usable resilience solution for scientific applications.
Within the context of three real applications we consider various combinations of algorithmic error detection, light-weight checkpoint restart to tolerate detected errors, as well as replication of pointers within data structures.
We focus on the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
Each application requires different resilience techniques, each of which can be configured in different ways.
We use extensive fault injection experiments to measure the performance/resilience properties of a large space of resilience configuration within each application.
For each configuration we measure the slowdown induced by the execution of resilience logic and roll-backs due to error detections, as well as the distribution of magnitudes of error in the application's output.
We then demonstrate how these measurements can be used to configure the resilience mechanisms used with each application to minimize overhead while meeting the user's accuracy requirements.
Since in practice users do not have a clear idea of the level of accuracy needed from an algorithm's results, we demonstrate a novel alternative method to specify accuracy requirements as the level noise in the application's input required to deviate the output by a given magnitude (i.e. similar to backward error analysis).
Finally, we extrapolate from the small-scale application runs considered in our fault injection campaign (the many runs required to cover the large space of resilience configurations limited the length of each run) to application runs with 1 million CPU hours via three different models of error propagation.
Our results demonstrate that when applications use all the relevant resilience methods, configured for maximum efficiency, the overhead of resilient execution on expected Exascale designs is low.

\greg{Need a few highlight with concrete numbers}
